Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-09-27 ms.Rnw
Fixed errors due to package update

In additional to probability of elimination estimates, we provide estimates for the cumulative number of infections under each vaccination scenario from February 2019 -- February 2024.
Notably, the median number of cumulative cholera infections under the no-vaccination scenario using Models 1 and 3 were $\Sexpr{prettyNum(round(as.numeric(mod1_V0_probs$cumInf['q50'])), big.mark = ",")}$ and $\Sexpr{prettyNum(round(as.numeric(mod3_V0_res$mod3_V0_probs$cumInf['q50'])), big.mark = ",")}$, respectively.
While there is remaining time during this projection period in which new cholera infections can be detected, up to this point our estimates are far more consistent with the observed number of reported cholera cases than the corresponding estimates from \cite{lee20}, which were approximately $400,000$ and $1,000,000$.

<<Create Model 1 VaccScen plots>>=
gg_m1_V0 <- ggplot() +
  geom_line(data = mod1_V0_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = mod1_V0_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
  theme(axis.title.x = element_blank()) +
  ggtitle("V0") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V1 <- ggplot() +
  geom_line(data = mod1_V1_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = mod1_V1_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
  theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V1") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V2 <- ggplot() +
  geom_line(data = mod1_V2_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = mod1_V2_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
    theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V2") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V3 <- ggplot() +
  geom_line(data = mod1_V3_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = mod1_V3_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
    theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V3") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V4 <- ggplot() +
  geom_line(data = mod1_V4_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = mod1_V4_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
    theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V4") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

@@

% \begin{figure}[ht]
% \centering
%
% mod1_cumInf <- data.frame(
%   'scenario' = paste0("V", 0:4),
%   'low' = c(
%     as.numeric(mod1_V0_probs$cumInf['q025']),
%     as.numeric(mod1_V1_probs$cumInf['q025']),
%     as.numeric(mod1_V2_probs$cumInf['q025']),
%     as.numeric(mod1_V3_probs$cumInf['q025']),
%     as.numeric(mod1_V4_probs$cumInf['q025'])
%   ),
%   'high' = c(
%     as.numeric(mod1_V0_probs$cumInf['q975']),
%     as.numeric(mod1_V1_probs$cumInf['q975']),
%     as.numeric(mod1_V2_probs$cumInf['q975']),
%     as.numeric(mod1_V3_probs$cumInf['q975']),
%     as.numeric(mod1_V4_probs$cumInf['q975'])
%   ),
%   'med' = c(
%     as.numeric(mod1_V0_probs$cumInf['q50']),
%     as.numeric(mod1_V1_probs$cumInf['q50']),
%     as.numeric(mod1_V2_probs$cumInf['q50']),
%     as.numeric(mod1_V3_probs$cumInf['q50']),
%     as.numeric(mod1_V4_probs$cumInf['q50'])
%   )
% )
%
% mod3_cumInf <- data.frame(
%   'scenario' = paste0("V", 0:4),
%   'low' = c(
%     as.numeric(mod3_V0_probs$cumInf['q025']),
%     as.numeric(mod3_V1_probs$cumInf['q025']),
%     as.numeric(mod3_V2_probs$cumInf['q025']),
%     as.numeric(mod3_V3_probs$cumInf['q025']),
%     as.numeric(mod3_V4_probs$cumInf['q025'])
%   ),
%   'high' = c(
%     as.numeric(mod3_V0_probs$cumInf['q975']),
%     as.numeric(mod3_V1_probs$cumInf['q975']),
%     as.numeric(mod3_V2_probs$cumInf['q975']),
%     as.numeric(mod3_V3_probs$cumInf['q975']),
%     as.numeric(mod3_V4_probs$cumInf['q975'])
%   ),
%   'med' = c(
%     as.numeric(mod3_V0_probs$cumInf['q50']),
%     as.numeric(mod3_V1_probs$cumInf['q50']),
%     as.numeric(mod3_V2_probs$cumInf['q50']),
%     as.numeric(mod3_V3_probs$cumInf['q50']),
%     as.numeric(mod3_V4_probs$cumInf['q50'])
%   )
% )
%
% mod1_cumInf$model <- "1"
% mod3_cumInf$model <- "3"
%
% all_cumInf <- rbind(mod1_cumInf, mod3_cumInf)
% all_cumInf$scenario <- factor(all_cumInf$scenario, levels = rev(c("V0", "V1", "V2", "V3", "V4")))
%
%
% ggplot(all_cumInf, aes(y = scenario)) +
%   geom_point(aes(x = med)) +
%   geom_linerange(aes(xmin = low, xmax = high)) +
%   scale_x_continuous(
%     labels = scales::unit_format(scale = 1e-6, accuracy = 1),
%     limits = c(0, 5e6)
%   ) +
%   xlab("Cumulative infections (million)") +
%   theme(axis.title.y = element_blank(),
%         axis.title.x = element_text(size = 10),
%         axis.text = element_text(size = 8)) +
%   facet_wrap(~model, ncol = 1,
%              labeller = as_labeller(
%                c("1" = "Model 1", "3" = "Model 3", "2" = "Model 2")
%              ))
%
% \caption{\label{fig:cumInfPlot}
% Simulated cumulative infections under various vaccination scenarios from February, 2019, to February, 2024.
% Point estimates (median cumulative incidence) and $95\%$ error bars are included.
% x-axis scale is intended to ease comparison with Figure~4 of \cite{lee20}.
% }
% \end{figure}

<<Compute Elimination Probs, echo=FALSE>>=
mod1_V0_probs$ElimTime$scenario = "V0"
mod1_V1_probs$ElimTime$scenario = "V1"
mod1_V2_probs$ElimTime$scenario = "V2"
mod1_V3_probs$ElimTime$scenario = "V3"
mod1_V4_probs$ElimTime$scenario = "V4"

mod1_probElims <- rbind(
  mod1_V0_probs$ElimTime,
  mod1_V1_probs$ElimTime,
  mod1_V2_probs$ElimTime,
  mod1_V3_probs$ElimTime,
  mod1_V4_probs$ElimTime
)

>>>>

In additional to probability of elimination estimates, we provide estimates for the cumulative number of infections under each vaccination scenario from February 2019 -- February 2024.
Notably, the median number of cumulative cholera infections under the no-vaccination scenario using Models 1 and 3 were $\Sexpr{prettyNum(round(as.numeric(h1_S0_results$mod1_V0_probs$cumInf['q50'])), big.mark = ",")}$ and $\Sexpr{prettyNum(round(as.numeric(mod3_V0_res$mod3_V0_probs$cumInf['q50'])), big.mark = ",")}$, respectively.
While there is remaining time during this projection period in which new cholera infections can be detected, up to this point our estimates are far more consistent with the observed number of reported cholera cases than the corresponding estimates from \cite{lee20}, which were approximately $400,000$ and $1,000,000$.

<<Create Model 1 VaccScen plots>>=
gg_m1_V0 <- ggplot() +
  geom_line(data = h1_S0_results$mod1_V0_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = h1_S0_results$mod1_V0_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
  theme(axis.title.x = element_blank()) +
  ggtitle("V0") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V1 <- ggplot() +
  geom_line(data = h1_S2_results$mod1_V1_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = h1_S2_results$mod1_V1_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
  theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V1") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V2 <- ggplot() +
  geom_line(data = h1_S4_results$mod1_V2_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = h1_S4_results$mod1_V2_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
    theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V2") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V3 <- ggplot() +
  geom_line(data = h1_S3_results$mod1_V3_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = h1_S3_results$mod1_V3_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
    theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V3") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

gg_m1_V4 <- ggplot() +
  geom_line(data = h1_S1_results$mod1_V4_sims, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(data = h1_S1_results$mod1_V4_sims, aes(x = date, ymin = q05 + 1, ymax = q95 + 1), alpha = 0.5) +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  ylab("Log reported cases") +
    theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  ggtitle("V4") +
  scale_x_date(date_labels = "'%y") +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)))

@@

<<Compute Elimination Probs, echo=FALSE>>=
h1_S0_results$mod1_V0_probs$ElimTime$scenario = "V0"
h1_S2_results$mod1_V1_probs$ElimTime$scenario = "V1"
h1_S4_results$mod1_V2_probs$ElimTime$scenario = "V2"
h1_S3_results$mod1_V3_probs$ElimTime$scenario = "V3"
h1_S1_results$mod1_V4_probs$ElimTime$scenario = "V4"

mod1_probElims <- rbind(
  h1_S0_results$mod1_V0_probs$ElimTime,
  h1_S2_results$mod1_V1_probs$ElimTime,
  h1_S4_results$mod1_V2_probs$ElimTime,
  h1_S3_results$mod1_V3_probs$ElimTime,
  h1_S1_results$mod1_V4_probs$ElimTime
)
